### 2020-09-28 paper/paper.tex
working on intersection and paper

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[table]{xcolor}
\usepackage{stmaryrd}
\usepackage{amsfonts}

% Code with syntax highlighting
\usepackage{fancyvrb}
\usepackage{listings}
\usepackage{textcomp}
\usepackage{placeins}

\theoremstyle{plain}% default
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem*{cor}{Corollary}
\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]
\newtheorem{exrc}[exmp]{Exercise}
\theoremstyle{remark}
\newtheorem*{comm}{Comment}
\newtheorem*{note}{Note}
\newtheorem{caso}{Case}

\def\C{\mathbb{C}}
\def\N{\mathbb{N}}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\K{\mathbb{K}}

\DeclarePairedDelimiter\sqmap{\llbracket}{\rrbracket}
\DeclarePairedDelimiter\spanset{\langle}{\rangle}

\newcommand{\diag}[1]{\text{diag}\left(#1\right)}
\newcommand{\mspan}[1]{\text{span}\left(#1\right)}
\newcommand{\dom}{\text{dom}}
\newcommand{\mker}[1]{\text{ker}\left(#1\right)}
\newcommand{\mrank}[1]{\text{rank}\left(#1\right)}

% matrix commands
\newcommand{\vvec}[2]{
	\begin{pmatrix}
		#1 \\ #2
	\end{pmatrix}
}
\newcommand{\vvvec}[3]{
	\begin{pmatrix}
		#1 \\ #2 \\ #3
	\end{pmatrix}
}
\newcommand{\vvvvec}[4]{
	\begin{pmatrix}
		#1 \\ #2 \\ #3 \\ #4
	\end{pmatrix}
}

% Code Listings
\definecolor{vgreen}{RGB}{104,180,104}
\definecolor{vblue}{RGB}{49,49,255}
\definecolor{vorange}{RGB}{255,143,102}
\lstset{
	basicstyle=\ttfamily,
	columns=fullflexible,
	keepspaces=true,
}

\title{Go Implementation of Up-To Techniques for Equivalence of Weighted Languages}

%\thanks{Use footnote for providing further
%information about author (webpage, alternative
%address)---\emph{not} for acknowledging funding agencies.}

\author{
  Alessandro Cheli \\
  Undergraduate Student \\
  Department of Computer Science \\
  Università di Pisa\\
  Pisa, PI 56127 \\
  \texttt{a.cheli6@studenti.unipi.it} \\
  %% examples of more authors
  %% \And
  %% Elias D.~Striatum \\
  %% Department of Electrical Engineering\\
  %% Mount-Sheikh University\\
  %% Santa Narimana, Levand \\
  %% \texttt{stariate@ee.mount-sheikh.edu} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle

\begin{abstract}
Weighted automata generalize non-deterministic automata by adding
a quantity expressing the weight (or probability) of the execution of each transition.
In this work we propose an implementation of two algorithms for computing language 
equivalence in finite state weighted automata (WAs). The first, a
linear partition refinement algorithm, calculates the largest linear weighted bisimulation
for any given LWA (Linear Weighted Automaton) through an iterative method, 
the second algorithm checks the language equivalence 
of two vectors (states) for a given weighted automata by using an additional
data structure representing a congruence relationship between states.
We then compare the two algorithms results to verify their correctness
on randomly generated automata samples, and provide some runtime statistics.

\end{abstract}

% keywords can be removed
\keywords{First keyword \and Second keyword \and More}

\section{Introduction}
\label{sec:intro}

In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.
For ease of implementation, we focus only on weighted automata over the field of real numbers 
$\R$.

\section{Preliminaries and Notation}
\label{sec:notation}

\begin{defn}
  A \textit{weighted automaton} over a field $\K$ and an alphabet $A$ is a triple 
  $(X,o,t)$ such that $X$ is a finite set of states, $t = (t_a : X \to K)_{a \in A}$
  is a set of transition functions indexed over the symbols of the alphabet $A$ and $o : X 
  \to \K$ is the output function. The transition functions $t_a$ are represented as
  $\K^{X \times X}$ matrices. $o \in \K^{1 \times X}$ is represented as a row vector.
  $t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
  vector $v  \in \K^{X \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
  dot product of the row vector $o$ with $v \in \K^{X \times 1}$.
  $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
  concatenation of a symbol $a$ to the word $w \in A^*$.
  A weighted language is a function $\psi: A^* \to \K$.
  A function mapping each state vector into its 
  accepted language, $\sqmap{\cdot}: \K^X \to K^{A^*}$ is defined as follows for every weighted automaton:

  \begin{equation*}
    \begin{aligned}
      \forall v \in \K^X, a \in A, w \in A^* \quad \quad
      \sqmap{v}(\epsilon) = o(v) \quad \quad
      \sqmap{v}(aw) = \sqmap{t_a(v)}(w)  
    \end{aligned}
  \end{equation*}

  Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called language equivalent, denoted with 
  $v_1 \sim v_2 $ if and only if $ \sqmap
  {v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
  than for vectors by assigning to each state $x \in X$ the corresponding unit vector 
  $e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, the language 
  of the automaton can be defined as $\sqmap{i}$.
\end{defn}

\begin{defn}
  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a subset of the 
  cartesian product of the sets. A relation is called \textit{homogeneous} or an \textit
  {endorelation} if it is a binary relation over $X$ and itself: $R \subseteq X \times
   X$. 
  In such case, it is simply called a binary relation over $X$.
  An \textit{equivalence relation} is a binary relation that is reflexive, symmetric and
  transitive. 
\end{defn}

An equivalence relation which is compatible with all the operations of
the algebraic structure on which it is defined on, is called a 
\textit{congruence relation}. Compatibility with the algebraic structure operations
means that algebraic operations applied on equivalent elements will still
yield equivalent elements.

\begin{defn}
  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
  $R'$ such that $R \subseteq R'$ 
\end{defn}

\begin{defn}
  \textbf{} \\

\end{defn}

\section{Algorithms}

\subsection{Forward Partition Refinement Algorithm for the Largest Weighted Bisimulation}
\label{sec:algo2}

The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}, & \quad & R_{i+1} = R_i \cap \bigcap_{a \in A} t_a(R_i)^{-1} \label{fwd} 
\end{eqnarray}
The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.1 of \cite{BONCHI201277}

\subsection{Backwards Partition Refinement Algorithm for the Largest Weighted Bisimulation}
\label{sec:algo2}

The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a(R_i)^t \label{back} 
\end{eqnarray}
The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.2 of \cite{BONCHI201277}

\section{Implementation}
\label{sec:impl}

The algorithms and data structures for this paper are implemented in the Go programming 
language. This implementation makes use of the \textit{Gonum} library for numerical 
computations. We only import the Gonum libraries for matrices and linear algebra 
and visual plotting of samples and functions.

Real numbers are implemented with double precision floating point numbers,  
known as the \texttt{float64} type in the Go programming language.

See Section \ref{sec:headings}.

\subsection{Input file format}

\subsection{Calculating the intersection of linear subspaces}

To compute algorithm (\ref{fwd}), we need an efficient way to calculate a basis 
of the intersection of an arbitrary number of linear subspaces of $\R^n$, given their 
spanning sets.

\begin{defn}
  \textbf{Algorithm for a basis of the intersection of subspaces} \\
  Defined in file \texttt{lin/intersect.go}.
  Let's consider two linear subspaces of 
  $\R^n$ defined with their spanning sets of column vectors: 
  $U = \spanset{u_1, u_2, \hdots, u_p}$ 
  and $W = \spanset{w_1, w_2, \hdots, w_k}$.
  We create the block matrix $A \in \R^{n \times (p+k)}$ defined as follows:

  \begin{equation*}
    \begin{aligned}
      a_{ij} = \begin{cases}
        (u_j)_i & \text{for  } j = 1, \hdots, p \quad i = 1, \hdots, n \\
        (w_{j-p})_i & \text{for  } j = p+1, \hdots, p+k \quad i = 1, \hdots, n \\
      \end{cases}
    \end{aligned}
  \end{equation*}

  The matrix will have the form
  \begin{equation*}
    A = \begin{bmatrix}
      \vvvec{}{u_1}{\,}, \hdots \vvvec{}{u_p}{\,}, 
      \vvvec{}{w_1}{\,}, \hdots \vvvec{}{w_k}{\,} 
    \end{bmatrix}
  \end{equation*}

  It follows that $\mker{A}$ is a basis for $U \cap W$. This can be done with an 
  arbitrary finite number of subspaces of $\R^n$, as long as their spanning set is 
  \textit{finite}.
\end{defn}

\begin{defn}
  \textbf{Algorithm for computing the nullspace} \\
  The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
  adapted from \cite{scipy/ranknullspace}. 
  Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:

  \begin{equation*}
    \begin{aligned}
      A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_{\min(m, n)}  } 
       & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
    \end{aligned}
  \end{equation*}

  Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_{\min(m,n)} \geq 0$.
  It follows that $\mrank{A}$ is equal to the number of nonzero singular values and
  a basis of the (right) null space of $A$ is the spanning set of the columns of V
  corresponding to singular values equal to zero. 
\end{defn}

\begin{exmp}
  The actual Go implementation is quite verbose, we can show a shorter Python 
  implementation of the algorithm to compute the nullspace, using the 
  \textit{SciPy} library \cite{scipy/ranknullspace}:

  \begin{lstlisting}[language=python]
def nullspace(A, atol=1e-13, rtol=0):
    A = np.atleast_2d(A)
    u, s, vh = svd(A)
    tol = max(atol, rtol * s[0])
    nnz = (s >= tol).sum()
    ns = vh[nnz:].conj().T
    return ns
    \end{lstlisting}
\end{exmp}

\subsubsection{Headings: third level}

\paragraph{Paragraph}

\section{Examples of citations, figures, tables, references}
\label{sec:others}

\cite{kour2014real,kour2014fast} and see \cite{hadash2018estimate}.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}

\subsection{Figures}
See Figure \ref{fig:fig1}. Here is how you add footnotes. \footnote{Sample of the first footnote.}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

\subsection{Tables}
See awesome Table~\ref{tab:table}.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

\bibliographystyle{unsrt}
\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
%%% and comment out the ``thebibliography'' section.

% 
% %%% Comment out this section when you \bibliography{references} is enabled.
% \begin{thebibliography}{1}
% 
% \bibitem{kour2014real}
% George Kour and Raid Saabne.
% \newblock Real-time segmentation of on-line handwritten arabic script.
% \newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
%   International Conference on}, pages 417--422. IEEE, 2014.
% 
% \bibitem{kour2014fast}
% George Kour and Raid Saabne.
% \newblock Fast classification of handwritten on-line arabic characters.
% \newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
%   International Conference of}, pages 312--318. IEEE, 2014.
% 
% \bibitem{hadash2018estimate}
% Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
%   Jacovi.
% \newblock Estimate and replace: A novel approach to integrating deep neural
%   networks with existing applications.
% \newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
% 
% \end{thebibliography}
% 
% 
\end{document}