### 2020-09-30 paper/notused.tex
preliminary definitions

\subsection{Calculating the intersection of linear subspaces}

To compute algorithm (\ref{fwd}), we need an efficient way to calculate a basis 
of the intersection of an arbitrary number of linear subspaces of $\R^n$, given their 
spanning sets.

\begin{defn}
  \textbf{Algorithm for a basis of the intersection of subspaces} \\
  Defined in file \texttt{lin/intersect.go}.
  Let's consider two linear subspaces of 
  $\R^n$ defined with their spanning sets of column vectors: 
  $U = \spanset{u_1, u_2, \hdots, u_p}$ 
  and $W = \spanset{w_1, w_2, \hdots, w_k}$.
  We create the block matrix $A \in \R^{n \times (p+k)}$ defined as follows:

  \begin{equation*}
    \begin{aligned}
      a_{ij} = \begin{cases}
        (u_j)_i & \text{for  } j = 1, \hdots, p \quad i = 1, \hdots, n \\
        (w_{j-p})_i & \text{for  } j = p+1, \hdots, p+k \quad i = 1, \hdots, n \\
      \end{cases}
    \end{aligned}
  \end{equation*}

  The matrix will have the form
  \begin{equation*}
    A = \begin{bmatrix}
      \vvvec{}{u_1}{\,}, \hdots \vvvec{}{u_p}{\,}, 
      \vvvec{}{w_1}{\,}, \hdots \vvvec{}{w_k}{\,} 
    \end{bmatrix}
  \end{equation*}

  It follows that $\mker{A}$ is a basis for $U \cap W$. This can be done with an 
  arbitrary finite number of subspaces of $\R^n$, as long as their spanning set is 
  \textit{finite}.
\end{defn}