### 2020-10-05 paper/sections/introduction.tex

\section{Introduction}
\label{sec:intro}
TODO da fare

Checking whether two nondeterministic finite automata (NFA) accept the same
language is important in many application domains such as compiler construction and model checking. 
Unfortunately, solving this problem is costly: it is
PSPACE-complete (TODO cita) 
% TODO cita

In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.

A well detailed example of the comparison of algorithms to compute 
language equivalence, precisely between \texttt{HKC} and an alternative
algorithm called the \textit{antichain algorithm} (\cite{de2006antichains}),
was published in 2017 \cite{fu2017equivalence}.