### 2020-10-05 paper/sections/implementation.tex

\section{Implementation}
\label{sec:impl}

The algorithms and data structures are implemented in the Go programming 
language. Real numbers are implemented with double precision floating point numbers,  
precisely of \texttt{float64} type.

This implementation makes use of the Gonum library, 
an excellent toolkit for high-performance numerical computations.
We only import the Gonum libraries for matrices, linear algebra 
and visual plotting.
Although not GPU accelerated, Gonum matrix operations are run on 
the CPU and accelerated with BLAS and LAPACK.

\subsection{Data Structures}
In this implementation, the data structure for representing weighted automata is a \texttt{struct} containing
the following attributes:
\begin{enumerate}
    \item An integer \texttt{Dim} representing the number of states $n$. 
    (note: states are finite and indexed on natural numbers $0, \hdots, n-1$)
    \item A slice of strings representing the \textit{alphabet} $A$.
    \item A map of strings as keys (the alphabet symbols) and dense $n \times n$ \texttt{float64} matrices as values,
    representing the set of transition functions.
    \item A dense \texttt{float64} vector, representing the linearization of the output function. 
    \item A \texttt{float64} value representing the tolerance to be used in numerical computations.
    \item An optionally \texttt{nil}, dense floating point matrix providing a basis for the orthogonal 
    complement of $\llwb$
\end{enumerate}
\begin{note}
    \textit{Slices} in Go are a convenient and efficient extension of the concept of arrays: 
    they provide an abstraction for indexed, variable length sequences of typed data, and 
    provide useful helper functions for creating, appending and selecting elements. 
\end{note}
The definition can be found in file \texttt{automata/automata.go}.
We then provide methods for reading an automaton from a text stream, applying transition
and output functions to vectors, and generating random automata with real and natural valued weights.

\subsection{Implementation of HKC}
Instead of creating dedicated structs,
we exploit the \texttt{mat.Dense} data type in Gonum to efficiently represent:
\begin{itemize}
    \item Sets of vectors with $n \times k$ dense matrices ($k$ is the number of vectors in the set)
    \item Pairs of vectors with $n \times 2$ dense matrices
    \item Sets and the "\texttt{todo}" stack of pairs with $n \times 2k$ dense matrices ($k$ is the number of pairs in the set or stack)
\end{itemize}

To increase efficiency of the methods for inclusion checking and insertion in sets of vectors,
one could keep the columns ordered (by vector norm) in the corresponding matrix.

To represent the congruence relation $R$, we introduce a \texttt{struct} containing:
\begin{itemize}
    \item A dense matrix \texttt{s} of size $n \times 2k$ containing the set of pairs in the relation
    \item A dense matrix \texttt{u} to represent the generating set $U_R$ for the congruence closure (see definition \ref{def:congclos}).
    \item Two integers representing the number of pairs in the set, and the size of vectors.
    \item A tolerance value to be used in equivalence checks. Best results were obtained with $10^{-14}$.
\end{itemize}

When adding a pair of vectors to the congruence relation, we extend the columns
of the matrix \texttt{s} with the pair $(v_1, v_2)$, if and only if $(v_1 - v_2)$ 
is not already in $U$. To check if the pair is in the congruence closure $c(R)$, 
we check if $(v_1 - v_2)$ is contained in U.

\captionof{figure}{Implementation of the \texttt{HKC}$(v_1,v_2)$ procedure}
\lstinputlisting[language=go]{../automata/hkc.go}

%TODO sistema

\subsection{Implementation of BPR}
Given a WA $L$, at the first step of \texttt{BPR}, we set $R_0 = o$, 
with $o$ being the dense vector representing the output function 
of $L$, as seen in \cite{BONCHI201277}.
For each step $i = 0, \hdots, j+1$, with $j \leq \dim(L)$, the implemented algorithm:
\begin{enumerate}
    \item For each $a \in A$ computes $t_a^T(R_i)$ through matrix multiplication
    \item Concatenates those matrices $t_a^T(R_i)$ to $R_i$ in a resulting matrix $G$ 
    \item Instead of checking linear independence of the columns through Gaussian elimination,
    it computes $R_{i+1}$ as the orthonormal basis of the column space of $G$, through singular 
    value decomposition.
\end{enumerate}

At the end of \texttt{BPR}, we store the basis for the orthogonal complement
of $\llwb$ as an attribute of the automaton. To check if two vectors $v_1 \llwb v_2$,
we check that $R_j^T(v_1 - v_2) = \vec{0}$, with a prefixed tolerance.

To compute a basis for $\llwb$, at the last step of the algorithm,
we would need to compute $R_j^0$.
If $V$ is a vector space and $W$ is a
subspace of $W$, the annihilator of $W$, respectively $W^0$ is 
a subspace of the space $V^*$ of linear functionals on $V$.
$W^0$ are the functionals that annihilate on $W$. Since 
we are working on subspaces of $\R^n$, we can directly compute 
the orthogonal complement in our implementation instead of the
annihilator.

\begin{prop}
  If $V$ is a finite dimensional vector space defined with an inner product
  $\langle \cdot , \cdot \rangle$ and $W$ is a subspace of $V$
  then the image of the annhilator $W^0$ through the linear 
  isomorphism $\varphi: V^* \to V$ induced by the inner product, 
  is the orthogonal of $W$ with respect to the said inner product.
\end{prop}

\begin{proof}
  Let $V$ be an inner product space over the field $\K$ with an inner product defined as
  $\langle \cdot , \cdot \rangle : V \times V \to \K$. 
  Every linear functional can be 
  represented with a vector. Let $\xi : V \to \K$ be a functional, 
  $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
  if $v$ represents $\xi$ we have that $(v, w)=\xi(w)=0$ for all $w \in W$. 
  We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
  If $v \in W^\perp$  
  then the functional $x \mapsto (v, x)$ cancels over $W$ 
  (by the definition of orthogonality).
\end{proof}

To compute the orthogonal complement of a vector subspace $W$, we
compute $W^\perp = \mker{A^T}$, where $A$ is the matrix whose column space 
is $W$. Proof is available in \cite{ila}.

\captionof{figure}{Implementation of Backwards Partition Refinement}
\lstinputlisting[language=go]{../automata/backwards.go}

\begin{note}
    \textbf{Applications of SVD} \\

    Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:

    \begin{equation*}
      \begin{aligned}
        A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_r  } 
         & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
      \end{aligned}
    \end{equation*}

    Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_r \geq 0$.
    It follows that $\mrank{A}$ is equal to the number of nonzero singular values, and
    as explained in \cite{svd}:

    \begin{enumerate}
      \item  $\mrank{A} = \mrank{\Sigma} = r$
      \item The column space of $A$ is spanned by the first $r$ columns of $U$.
      \item The null space of $A$ is spanned by the last $n − r$ columns of $V$.
      \item The row space of $A$ is spanned by the first $r$ columns of $V$.
      \item The null space of $A^T$ is spanned by the last $m − r$ columns of $U$.
    \end{enumerate}

    Of our interest, are only the computation of the null space and column space.
    The implementation 
    can be found in files \texttt{lin/colspace.go} and \texttt{lin/nullspace.go}.
  \end{note}