### 2020-10-05 paper/sections/runtime.tex

\section{Runtime Results}

TODO da fare e sistemare.

The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. 
For 3 test runs on 100000 automata composed of 4 states, 2 alphabet symbols and transition
matrices with weights on natural numbers $\leq 3$, with a tolerance of 10e-14,
we found a non empty kernel of $\llwb$ for an average of 1718 automata.

%TODO
(TODO spiega: più aumentano i simboli/stati più si riduce la probabilità di 
trovare $\llwb$ non vuoto?? investiga, trova un modo di generare automi con LLWB 
non vuoto)

By generating 1000 random couples of language equivalent vectors $(v_1, v_2)$ such that
$(v_1 - v_2) \in \mker{\llwb}$ and 1000 couples of completely random vectors
$(w_1, w_2)$, for each automata, we verified with 100\% confidence that:

\begin{center}
    $v_1 \; R_j \; v_2 \iff$ \texttt{HKC}$(v_1, v_2) =$ \texttt{false}
\end{center}

Where $R_j$ is the largest linear weighted bisimulation $\llwb$ computed by \texttt{BPR}.

%TODO

Sample output (19992 KBs in memory):
\begin{verbatim}
    $ time go run main.go 
    tests completed            100000...

    1000 samples per automaton
    100000 automata
    LLWP is not empty for 1726 automata
    language equivalence HKC = BPR for 3452000 samples
    BPR equivalence verified for 1726000 samples
    HKC equivalence verified for 1726000 samples
    confidence is 100%

    real    0m21.322s
    user    0m23.614s
    sys     0m0.961s
\end{verbatim}