### 2020-10-11 paper/sections/runtime.tex
added tolerance tests, added comments to code. 0 problems in code

\section{Runtime Results}

%TODO da fare e sistemare.

%\subsection{Tolerance and F1 Score}
We the studied of the influence 
of the tolerance value over the correctness of the algorithms results.
Tests took place on randomly generated automata. A basis of  $\llwb$ was 
computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
as an uniformly distributed random linear combination of the vectors in the spanning set of $\llwb$. 
The same quantity of vector pairs was generated randomly with an uniform distribution.
BPREquivalence and HKC procedures were then executed with those pairs in input.

By defining pairs of vectors as

\begin{itemize}
    \item \textit{true positives} (TP) if reported as language equivalent by both 
    procedures
    \item \textit{true negatives} (TN) if reported as not language equivalent by both 
    procedures
    \item \textit{false positives} (FP) if reported as language equivalent by HKC but not by BPR
    \item \textit{false negatives} (FN) if reported as  language equivalent by BPR but not by HKC
\end{itemize}

We have introduced \textit{accuracy, precision, recall} and \textit{F1 score} over 
the tests results:

\begin{equation*}
    \begin{aligned}
        \text{Accuracy} = \dfrac{TP + TN}{TP + TN + FP + FN} & \quad & 
        \text{Recall} = \dfrac{TP}{TP + FN} \\ && \\
        \text{Precision} = \dfrac{TP}{TP + FP} & \quad &
        \text{F1} = 2 * \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
    \end{aligned}
\end{equation*}

We have then collected 
the F1 score in relation to varying tolerance values.
The plot is shown in Figure \ref{fig:f1}.  
The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. Executed sequentially, F1 score tests took 2m26.19s
Parallelized with a worker pool, the tests took 1m13.12s.

% //TODO: condizionamento BPR
TODO: BPR conditioning

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-06.pdf}
    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-13.pdf}
    \caption{F1 score over tolerance tests, 2000 vector pairs tested per automaton.}
    \label{fig:f1}
\end{figure}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-16.pdf}
    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-20.pdf}
    \caption{F1 score over tolerance tests, 2000 vector pairs tested per automaton.}
    \label{fig:f1-2}
\end{figure}

TODO empty LLWB chance, memory and time plots, varying automata size parameters